What Can Stack Traces Tell us about of the use Exceptions in Java?

### 2014-07-07 android-stacks.tex
merging introductions - step 3
Giving the introduction some love

\subsection{Georgios's and Roberta Merged Introduction}

Modern applications have to cope with an increasing number of abnormal
computation states that arise as a consequence of faults in the application
itself (e.g., access of null references), noisy user inputs or faults in
underlying middleware or hardware. The exception handling
mechanism~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions. When an exception is thrown it can be caught inside the application (by the program's exception handling code) and the the application has the opportunity to assess the conditions that caused the fault and either recover from them or abort. 
If the exception that remains uncaught (is not caught by an exception handler) it
leads to a program crash -  uncaught exceptions ~\cite{jo2004uncaught} are one 
of the main causes of application crashes.

 In most programming languages with embedded exception handling mechanism, when a crash happens usually the language runtime generates debugging information to help
the programmer identify and fix the cause of the crash. The debugging
information is typical a stack trace, which includes an explanatory message and
the execution stack frame when the exception occurred. In Java, stack trace
consists of an ordered list of methods that were active on the call stack before
the exception occurred.

Although the exception handling mechanism has been the target of several studies
(e.g.~\cite{miller1997issues,Robil00,shah2010understanding,
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
the exception handling code is often generally poorly understood and the least
tested part of software systems~\cite{coelho2011unveiling}.

%Similarily to source
%code, 
%TODO GEORGIOUS I DID NOT UNDERESTA.ND THE IDEA OF THE PREVIOUS SENTENCE

Such studies  perform the analysis of exception handling code based  either on static
(source code-based) analysis or dynamic (runtime-based) testing methods. Neither of such methods, however,
can explore the full execution space of a program and therefore they both fail
to account for exceptions implicitly signaled by rutime environment (e.g. resource depletion) or
dynamically loaded code. Exception stack traces, when available, enable the
post-mortem analysis of exception handling behaviours of programs and
concequently the evaluation of exception handling mechanisms in programming
languages.

The availability of exception stack traces until recently was scarce, while the
available data were only focused on few large scale projects.
Along the recent years we have witnessed a rapid increase in the number of open source 
projects that make their control version and issue tracking systems publicly 
available in repository hosting sites such as GitHub.

The concentration of open source projects and the existence of
initiatives that mine such information, such as GHTorrent~\cite{Gousi13},
 allow the information about projects to be visible within and across projects and
 make a full population studies possible.

In this work, we perform the first post mortem analysis of exception handling in
Java.  Using a custom tool called ExceptionMiner, which we develop specifically
for this study, we mine stack traces from issues across all Java projects on
GitHub. Overall, we analyzed 356,057 issues from 16,836 projects, from which
28,800 stack traces were extracted and used as a source of information for our
analysis. The stack trace analysis was augmented by additional information
extracted using byte code and source code analysis done on a carefully selected
subset of our initial sample (482 Android projects). Manual inspection was also 
used in this study to leverage the understanding of stack traces and support 
further discussions and insights.

To guide this exploration we compiled general 
guidelines on how to use exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
we investigate what stack traces can tell us about the adherence to such
practices in Java programs. Some outcomes were consistently detected through
our large scale analysis of exception stack traces in the wild, such as:

>>>>

\subsection{Georgios's and Roberta's Merged Introduction}

Modern applications have to cope with an increasing number of abnormal
computation states that arise as a consequence of faults in the application
itself (e.g., access of null references), noisy user inputs or faults in
underlying middleware or hardware. The exception handling
mechanism~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions. When an exception is
thrown it can be caught inside the application (by the program's exception
handling code) and the application has the opportunity to assess the conditions
that caused the fault and either recover from them or abort. If the exception
that remains uncaught, the program crashes. Uncaught exceptions are one of the
main causes of application crashes~\cite{jo2004uncaught}.

In most programming languages featuring exception handling, when a crash
happens, the language runtime generates debugging information to help the
programmer identify the cause of the crash. The debugging information is
typically a stack trace, which includes an explanatory message and the execution
stack frame when the exception occurred.
%In Java, stack trace consists of an ordered list of methods that were active on
%the call stack before the exception occurred.

Although the exception handling mechanism has been the target of several studies
(e.g.~\cite{miller1997issues,Robil00,shah2010understanding,
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
the exception handling code is generally poorly understood and often the least
tested part of software systems~\cite{coelho2011unveiling}.  To analyze the
exception handling code, current studies use either static (source code-based)
analysis or dynamic (runtime-based) testing methods. Neither method can explore
the full execution space of a program and therefore they both fail to account
for exceptions implicitly signaled by the runtime environment (e.g. due to
resource depletion) or dynamically loaded code. Exception stack traces, when
available, enable the post-mortem analysis of exception handling behaviours of
programs and consequently the evaluation of exception handling mechanisms in
programming languages.

The availability of exception stack traces until recently was scarce, while the
available data concerned a few large scale projects (e.g. the Mozilla
crash report
dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Along the recent years, we have witnessed a rapid increase in the number of open
source projects that make their control version and issue tracking systems
publicly available in repository hosting sites such as GitHub. The
concentration of open source projects and the existence of initiatives that mine
such information, such as GHTorrent~\cite{Gousi13}, allow the information about
projects to be visible within and across projects and make a full population
studies possible.

In this work, we perform the first post mortem analysis of exception handling in
Java. Using a custom tool called ExceptionMiner, which we develop specifically
for this study, we mine stack traces from issues across all Java projects on
GitHub. Overall, we analyzed 356,057 issues from 16,836 projects, from which
28,800 stack traces were extracted and used as a source of information for our
analysis. The stack trace analysis was augmented by additional information
extracted using byte code and source code analysis done on a carefully selected
subset of our initial sample (482 Android projects). Manual inspection was
used to leverage the understanding of stack traces and support
further discussions and insights.
To guide our analysis, we compiled a set of best practices on how to use
exceptions by composing relevant suggestions by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective}.  The
comparison of those with the mined exceptions revealed several interesting
findings, such as:
